Customizing Software by Direct Manipulation of Structured Data

### 2020-05-21 paper.tex
compile

\hypertarget{the-programming-cliff}{%
\subsubsection{The programming cliff}\label{the-programming-cliff}}

@@

It is a perfectly unreasonable choice to choose imperative programming
as the model for building customizations. Imperative programming
provides a high ceiling for possible customizations, and the model of a
sequence of commands is a natural fit for automation workflows which
simulate a series of steps taken by the user. However, there is a
serious drawback to this approach.

MacLean et al \citep{maclean1990} describe an ideal for user-tailorable
systems: a ``gentle slope'' from using to customizing, where small
incremental increases in skill lead to corresponding increments of
customization power. Requiring users to learn programming for
customization creates an abrupt ``cliff,'' where a significant learning
investment is required even to implement simple customizations. Another
goal of MacLean et al is that ``it should be as easy to change the
environment as it is to use it,'' at least for some subset of changes.
In scripting languages, even user-friendly ones like Applescript or
Chickenfoot, the experience of customization does not remotely resemble
the experience of use, so these systems can't meet this goal.

\hypertarget{a-gentler-slope}{%
\subsubsection{A gentler slope}\label{a-gentler-slope}}

>>>>

@@

It is reasonable to choose imperative programming as the model for
building customizations. Turing-Complete programming provides a high
ceiling for possible customizations, and the a sequence of commands is a
natural fit for automations which simulate a series of steps taken by
the user.

However, there is a serious drawback to this approach. MacLean et al
\citep{maclean1990} describe an ideal for user-tailorable systems: a
``gentle slope'' from using to customizing, where small incremental
increases in skill lead to corresponding increments of customization
power. Requiring users to learn programming for customization creates an
abrupt ``cliff,'' where a significant learning investment is required
even to implement simple customizations. Another goal of MacLean et al
is that ``it should be as easy to change the environment as it is to use
it,'' at least for some subset of changes. In scripting languages, even
user-friendly ones like Applescript or Chickenfoot, the experience of
customization does not remotely resemble the experience of use, so these
systems can't meet this goal.
